---
title: "Practical_Day_4"
format: html
editor: visual
---

## Practical Day 4

Setting Up

```{r}
# Install tidyverse
if (!requireNamespace("tidyverse", quietly = TRUE)) {
  install.packages("tidyverse")
}
library(tidyverse)

# Install and load nycflights13 for flight data
if (!requireNamespace("nycflights13", quietly = TRUE)) {
  install.packages("nycflights13")
}
if (!requireNamespace("remotes", quietly = TRUE)) {
  install.packages("remotes")
}
remotes::install_github("SATVILab/UtilsDataRSV")
library(nycflights13)
library(tidyr)
library(dplyr)
```

## Question 2

Displaying the Flights dataset

```{r}
#| echo: false
data_nycflights13 <- as_tibble(flights)
dplyr::glimpse(data_nycflights13)
UtilsDataRSV::view_cols(data_nycflights13)
##'print(data_nycflights13, width = Inf)
```

Rewriting the code

```{r}
flight1 <- flights%>%filter(month == 1)

dist_tbl <- flight1%>%group_by(carrier)%>%summarise(mean_distance = mean(distance, na.rm = TRUE),sd_distance = sd(distance, na.rm = TRUE))
view(dist_tbl)
```

# Question 3

The distance is the same for all routes with carrier identifier being AS. That is why the standard deviation is 0. The is only one route recorded with carrier identifier being OS and there iis

```{r}
AS_view <- flight1%>%filter(carrier == "AS")
AS_view 
OO_view <- flight1%>%filter(carrier == "OO")
OO_view
```

# Question 4

```{r}

flights_r <- flights |> group_by(month, carrier)  |> summarise(avg_delay = mean(dep_delay, na.rm = TRUE))
flights_r
flights_r |>  pivot_wider(names_from = carrier, values_from = avg_delay)

```

## Question 5
The proportion of the flights that were delayed and arrived on time is 10.52%.
```{r}
Number_flights <- flights |> filter(dep_delay >0, arr_delay<= 0) |> summarise(proportion = sum(dep_delay >0 & arr_delay <= 0, na.rm = TRUE))
Proportion = (Number_flights/nrow(flights)) * 100
Proportion

```

## Question 6
6.1 
```{r}
routes_ <- flights |> group_by(origin, dest) |> distinct(carrier) |> summarise(count = n(), .groups = "drop")
routes_g1 <- routes_ |> filter( count > 1)
```
6.2
